---
title: "ggscribe"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ggscribe}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

First, let's create a pretty base plot to work with.

ggscribe annotation functions generally require:

* a globally set theme with panel dimensions
* cartesian space with `clipping = "off"`

```{r setup, message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
library(stringr)

set_theme(
  ggrefine::theme_grey(
    panel_heights = rep(unit(50, "mm"), 100),
    panel_widths = rep(unit(75, "mm"), 100),
  )
)

p <- ggplot2::mpg |>
  mutate(drv = case_when(
    drv == "4" ~ "4-wheel",
    drv == "f" ~ "Front",
    drv == "r" ~ "Rear",
  )
  ) |>
  ggplot(aes(x = displ, y = hwy, fill = drv, colour = drv)) +
  geom_point() +
  scale_fill_discrete(palette = jumble::jumble) +
  scale_colour_discrete(palette = blends::multiply(ggplot2::get_theme()$palette.fill.discrete)) +
  labs(
    title = "Highway fuel economy",
    subtitle = "By displacement and drive train\n",
  ) +
  coord_cartesian(clip = "off") + #required to annotate outside the panel
  ggrefine::refine_modern()

p
```

Now let's demo some uses of ggscribe. 

1). A custom axis line.

```{r}
p +
  ggscribe::annotate_axis_line(
    position = "bottom",
    xmin = 2,
    xmax = 7,
    element_to = "transparent",
  ) 
```

2). Annotated labels along the top of the panel. Note how a different `hjust` is used for each label.

```{r}
p +
  ggscribe::annotate_axis_text(
    position = "top",
    x = I(0),
    label = "Low",
    tick_length = rel(0),
    hjust = 0,
  ) +
  ggscribe::annotate_axis_text(
    position = "top",
    x = I(0.5),
    label = "Normal",
    tick_length = rel(0),
    hjust = 0.5,
  ) +
  ggscribe::annotate_axis_text(
    position = "top",
    x = I(1),
    label = "High",
    tick_length = rel(0),
    hjust = 1,
  ) +
  labs(subtitle = "By displacement and drive train\n\n",)
```

3). A shaded band and text along the right hand side of the panel.

```{r}
p +
  ggscribe::annotate_panel_shade(ymax = 20, fill = flexoki::flexoki$red["red200"]) +
  ggscribe::annotate_axis_text(
    position = "right",
    y = c(20),
  ) +
  ggscribe::annotate_axis_ticks(
    position = "right",
    y = c(20),
  ) +
  ggscribe::annotate_axis_text(
    position = "right",
    y = (20 + (min(mpg$hwy) - 0.05 * diff(range(mpg$hwy)))) / 2,
    label = "Inefficient",
    element_to = "transparent",
  )
```

4). A shaded band and text along the right hand side of the panel - and positioned such that we need to create more space for the label.

```{r}
p +
  scale_y_continuous(
    sec.axis = dup_axis(
      labels = \(x) paste0(
        str_sub(x, 0, 0), #Create space for the word 'Efficient'
        str_flatten(rep(" ", times = str_length("Efficient") + 2))),
      name = NULL,
    )
  ) +
  ggscribe::annotate_panel_shade(ymin = 30, fill = flexoki::flexoki$blue["blue200"]) +
  ggscribe::annotate_axis_text(
    position = "right",
    y = c(30),
  ) +
  ggscribe::annotate_axis_ticks(
    position = "right",
    y = c(30),
  ) +
  ggscribe::annotate_axis_text(
    position = "right",
    y = (30 + (max(mpg$hwy) + 0.05 * diff(range(mpg$hwy)))) / 2,
    label = "Efficient",
  )
```

5). Mobile friendly y axis text.

```{r}
p +
  ggscribe::annotate_axis_text(
    y = c(20, 30, 40),
    element_to = "blank",
    tick_length = rel(-1),
    hjust = 0,
    vjust = -0.5,
  ) +
  ggscribe::annotate_axis_text(
    position = "top",
    x = I(0),
    label = "hwy",
    tick_length = rel(0),
    hjust = 0,
  ) +
  labs(
    subtitle = "By displacement and drive train\n\n",
    y = NULL,
  )
```

6). Text coloured dark or light depending on the fill.

```{r}
ggwidth::set_equiwidth(equiwidth = 1.75)

mtcars |>
  count(cyl, am) |>
  mutate(
    am = if_else(am == 0, "Automatic", "Manual"),
    cyl = as.factor(cyl)
  ) |>
  ggplot(aes(x = am, y = n, colour = cyl, fill = cyl, label = n)) +
  geom_col(
    position = position_dodge2(preserve = "single"),
    width = ggwidth::get_width(n = 2, n_dodge = 3),
  ) +
  scale_fill_discrete(palette = jumble::jumble) +
  scale_colour_discrete(palette = blends::multiply(jumble::jumble)) +
  geom_text(
    mapping = ggscribe::aes_contrast(), # or aes(!!!ggscribe::aes_contrast()),
    position = position_dodge2(
      width = ggwidth::get_width(n = 2, n_dodge = 3),
      preserve = "single"),
    vjust = 1.33,
    show.legend = FALSE,
  ) +
  scale_y_continuous(expand = expansion(c(0, 0.05))) +
  ggrefine::refine_modern(x_type = "discrete")
```

