---
title: "ggscribe"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with ggscribe}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`ggscribe` extends ggplot2 annotation capabilities. 

Note: 

* `annotate_axis_ticks` and `annotate_axis_text` require (1) a globally set theme with explicit panel dimensions and (2) `coord_cartesian(clip = "off")`
* `annotate_panel_shade` should be before geoms.

In general, use the relevant scale function and breaks argument to annotate the axis unless there is a reason not to.

```{r base-setup, message=FALSE}
library(tidyverse)
library(ggscribe)

set_theme(
  ggrefine::theme_grey(
    panel_heights = rep(unit(50, "mm"), 100),
    panel_widths  = rep(unit(75, "mm"), 100),
  )
)

p <- ggplot2::mpg |>
  mutate(drv = case_when(
    drv == "4" ~ "4-wheel",
    drv == "f" ~ "Front",
    drv == "r" ~ "Rear",
  )) |>
  ggplot(aes(x = displ, y = hwy)) +
  coord_cartesian(clip = "off") +
  scale_fill_discrete(palette = jumble::jumble) +
  scale_colour_discrete(palette = blends::multiply(ggplot2::get_theme()$palette.fill.discrete)) +
  labs(title = "Highway fuel economy", subtitle = "By displacement\n") +
  ggrefine::refine_modern()
```

Annotate a axis line.

```{r axis-line}
p +
  geom_point() +
  annotate_axis_line(
    position    = "bottom",
    xmin        = 2,
    xmax        = 7,
    element_to  = "transparent",
  )
```

Annotate a threshold without affecting gridlines etc.

```{r}
p +
  ggrefine::theme_grey() +
  annotate_axis_line(x = 2.33, linetype = "dashed") +
  annotate_axis_text(x = 2.33, label = "Threshold", tick_length = rel(0), position = "top") +
  geom_point() +
  labs(subtitle = "By displacement\n\n") +
  geom_point() +
  annotate_axis_line(y = 32, linetype = "dashed") +
  scale_y_continuous(
    sec.axis = dup_axis_text(
      breaks = 32, 
      labels = "Threshold",
    )
  )
```




Annotate axis text with different hjusts.

```{r axis-text-top}
p +
  geom_point() +
  annotate_axis_text(position = "top", x = I(0),   label = "Low",    tick_length = rel(0), hjust = 0) +
  annotate_axis_text(position = "top", x = I(0.5), label = "Normal", tick_length = rel(0), hjust = 0.5) +
  annotate_axis_text(position = "top", x = I(1),   label = "High",   tick_length = rel(0), hjust = 1) +
  labs(subtitle = "By displacement\n\n")
```

```{r axis-text-ticks}
p +
  geom_point() +
  annotate_axis_text(x = I(0.5), label = "Displacement", tick_length = rel(6)) +
  annotate_axis_text(x = I(0),   label = "Low",           tick_length = rel(6), hjust = 0) +
  annotate_axis_text(x = I(1),   label = "High",          tick_length = rel(6), hjust = 1) +
  labs(x = "")
```

Annotate axis text with space automatically added.

```{r panel-shade}
p +
  annotate_panel_shade(ymin = 32) +
  geom_point(aes(fill = drv, colour = drv)) +
  scale_y_continuous(
    sec.axis = dup_axis_text(
      breaks = \(x) mean(c(x[2], 32)),
      labels = "Inefficient",
    )
  )
```

Annotate with a bracket for a label.

```{r panel-shade-bracket}
p +
  annotate_panel_shade(ymin = 32) +
  geom_point(aes(fill = drv, colour = drv)) +
  scale_y_continuous(
    sec.axis = dup_axis_text(
      breaks = \(x) mean(c(x[2], 32)),
      labels = "Inefficient",
    )
  ) +
  annotate_axis_line(position = "right", ymin = 32,       ymax = Inf) +
  annotate_axis_ticks(position = "right", y   = c(32, Inf))
```

Annotate with no space automatically added.

```{r panel-shade-inline}
p +
  annotate_panel_shade(ymax = 20, fill = flexoki::flexoki$red["red200"]) +
  geom_point(aes(fill = drv, colour = drv)) +
  annotate_axis_text(
    position   = "right",
    y          = (20 + (min(mpg$hwy) - 0.05 * diff(range(mpg$hwy)))) / 2,
    label      = "Inefficient",
    element_to = "transparent",
  )
```

Add in a mobile-friendly y axis.

```{r mobile-y}
p +
  geom_point() +
  annotate_axis_text(
    y           = c(20, 30, 40),
    element_to  = "blank",
    tick_length = rel(-1),
    hjust       = 0,
    vjust       = -0.5,
  ) +
  annotate_axis_text(
    position    = "top",
    x           = I(0),
    label       = "hwy",
    tick_length = rel(0),
    hjust       = 0,
  ) +
  labs(subtitle = "By displacement\n\n", y = NULL)
```

Add a aesthetic that determines colour based on the fill.

```{r contrast-text}
ggwidth::set_equiwidth(equiwidth = 1.75)

mtcars |>
  count(cyl, am) |>
  mutate(
    am  = if_else(am == 0, "Automatic", "Manual"),
    cyl = as.factor(cyl)
  ) |>
  ggplot(aes(x = am, y = n, colour = cyl, fill = cyl, label = n)) +
  geom_col(
    position = position_dodge2(preserve = "single"),
    width    = ggwidth::get_width(n = 2, n_dodge = 3),
  ) +
  scale_fill_discrete(palette = jumble::jumble) +
  scale_colour_discrete(palette = blends::multiply(jumble::jumble)) +
  geom_text(
    mapping  = aes_contrast(),
    position = position_dodge2(
      width    = ggwidth::get_width(n = 2, n_dodge = 3),
      preserve = "single"
    ),
    vjust       = 1.33,
    show.legend = FALSE,
  ) +
  scale_y_continuous(expand = expansion(c(0, 0.05))) +
  ggrefine::refine_modern(x_type = "discrete")
```
