---
title: "ggscribe"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with ggscribe}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`ggscribe` extends ggplot2 annotation capabilities. 

Note: 

* `annotate_axis_ticks`, `annotate_axis_text` and `annotate_axis_bracket` require (1) a globally set theme with explicit panel dimensions and (2) `coord_cartesian(clip = "off")`
* `annotate_panel_shade` should be before geoms.

In general, use the relevant scale function and breaks argument to annotate the axis unless there is a reason not to.

```{r base-setup, message=FALSE}
library(ggplot2)
library(dplyr)

set_theme(
  ggrefine::theme_grey(
    panel_heights = rep(unit(50, "mm"), 100),
    panel_widths  = rep(unit(75, "mm"), 100),
  )
)

p <- ggplot2::mpg |>
  mutate(drv = case_when(
    drv == "4" ~ "4-wheel",
    drv == "f" ~ "Front",
    drv == "r" ~ "Rear",
  )) |>
  ggplot(aes(x = displ, y = hwy)) +
  coord_cartesian(clip = "off") +
  scale_fill_discrete(palette = jumble::jumble) +
  scale_colour_discrete(palette = blends::multiply(ggplot2::get_theme()$palette.fill.discrete)) +
  labs(title = "Highway fuel economy", subtitle = "By displacement\n")
```

Add reference lines and labels.

```{r}
p +
  ggscribe::annotate_reference_line(xintercept = 2.4) +
  ggscribe::annotate_reference_line(yintercept = 32)  +
  geom_point(aes(colour = drv, fill = drv)) +
  scale_x_continuous(
    sec.axis = ggscribe::dup_axis_text(
      breaks = 2.4,
      labels = "Threshold",
    )
  ) +
  scale_y_continuous(
    sec.axis = ggscribe::dup_axis_text(
      breaks = 32, 
      labels = "Threshold",
    ) 
  ) 
```


